Frequent Payments as a Commitment Device?

### 2020-05-30 main.tex
update repo structure

\begin{table}[H]
\caption{Intermediate outcomes}
\label{mechanisms}
\begin{center}
\scriptsize{\input{./Tables/mechanism.tex}}
\end{center}
 \scriptsize
This table explores treatment effects in ``intermediate variables''. Each column represents a different dependent variable in an OLS regression. And each Panel represents a different treatment arm-control comparison. All regressions include fixed effects for branch, day of week fixed effects, number of pawns at the time of pawning this particular one, number of different treatment arms experienced before pawning this particular one. $R^2$ are not reported, they were less than 0.04 for all but the last column \hl{which had $R^2$'s close to 0.6}. The last row shows the mean of the dependent variable for the control group. Dependent variables are as follows: Column (1) measures the number of days elapsed from the pawning date to the first payment done. Column (2) records the number of payments done \hl{within 150 days} of the pawn. Column (3) is the average size of the payments for each client in pesos. Column (4) is self-reported cost to get to the branch plus the imputed loss of a whole day of salary (using the minimum wage in Mexico) multiplied by the number of visits to the branch. Column (5) measures the number of days it took to recover the pawn, conditional on recovering. The dependent variable in Column (6) is a dummy =1 if the client paid any positive amount within 150 days of the pawn. For Columns (7) and (8) the dependent variable it is the sum of total payments done within 150 days of the pawn divided by the size of the loan, but column (8) includes as regressors a dummy=1 if the client recovered her pawn, and its interaction with treatment.

\textit{Do file: } \texttt{mechanisms.do}
\end{table}

@@

\begin{figure}[H]
     \caption{Experiment description}
    \label{exp_description}
    \begin{center}
    \begin{subfigure}{.50\textwidth}
      \caption{Timeline of the experiment}
        \centering
        \includegraphics[width=\textwidth]{Figuras/timeline_suc_exp_extended.pdf}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
    \caption{Experiment arms}
       \centering
      \includegraphics[width=\textwidth]{Figuras/exp_arms.PNG}
    \end{subfigure}
    \end{center}
         \scriptsize Panel (a) displays in black arrows the period covered by our administrative data: from \hl{XXX} (\hl{XXX} days before the experiment began) to \hl{XXX}. The gray lines arrows correspond to the dates of implementation of the experiment. We finished the experiment earlier in branches 1 and 2 because we realized that we would have enough sample size with the remaining branches and save on operational costs. Panel (b) shows our 5 treatment arms: control (status quo contract), no choice with fee, no choice with promise, choice among both contracts where the commitment contract has a fee, and choice among both contracts where the commitment contract involves only a promise. Within each arm it displays the number of observations. And  for the choice groups it splits the sample size into those who chose the commitment contract (shaded blue triangle) and those that did not (un-shaded triangle).
      \textit{Do file: }  \texttt{timeline\_suc\_exp.do}
\end{figure}

@@

    \begin{subfigure}{0.45\textwidth}
        \caption{Financial cost -- Promise-forcing}
        \centering
        \includegraphics[width=\textwidth]{Figuras/fc_te_pro_3.pdf}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
        \caption{Losing Pawn -- Promise-forcing}
        \centering
        \includegraphics[width=\textwidth]{Figuras/def_te_pro_3.pdf}
    \end{subfigure}
    \end{center}
             \footnotesize \textit{Notes: } 
      \footnotesize{ } \textit{Do file: }  \texttt{fc\_te\_choice\_dec.do, def\_te\_choice\_dec.do, fc\_quantilereg\_choice\_dec.do}
\end{figure}

@@

\begin{figure}[H]
        \caption{ECDF of Financial Cost}
    \label{ecdf_fc}
    \begin{center}
        \centering
        \includegraphics[width=0.5\textwidth]{Figuras/cdf_fc_pro_2.pdf}
    \end{center}
    \footnotesize \textit{Notes: } 
     \footnotesize{ \textit{Do file: }  \texttt{hist\_fc.do}}
\end{figure}

>>>>

\begin{table}[H]
\caption{Intermediate outcomes}
\label{mechanisms}
\begin{center}
\scriptsize{\input{./Tables/mechanism.tex}}
\end{center}
 \scriptsize
This table explores treatment effects in ``intermediate variables''. Each column represents a different dependent variable in an OLS regression, and each panel represents a different treatment arm-control comparison. All regressions include fixed effects for branch, day of week fixed effects, number of pawns at the time of pawning this particular one, number of different treatment arms experienced before pawning this particular one. $R^2$ are not reported, they were less than 0.04 for all but the last column \hl{which had $R^2$'s close to 0.6}. The last row shows the mean of the dependent variable for the control group. Dependent variables are as follows: Column (1) measures the number of days elapsed from the pawning date to the first payment done. Column (2) records the number of payments done \hl{within 150 days} of the pawn. Column (3) is the average size of the payments for each client in pesos. Column (4) is self-reported cost to get to the branch plus the imputed loss of a whole day of salary (using the minimum wage in Mexico) multiplied by the number of visits to the branch. Column (5) measures the number of days it took to recover the pawn, conditional on recovering. The dependent variable in Column (6) is a dummy =1 if the client paid any positive amount within 150 days of the pawn. For Columns (7) and (8) the dependent variable it is the sum of total payments done within 150 days of the pawn divided by the size of the loan, but column (8) includes as regressors a dummy=1 if the client recovered her pawn, and its interaction with treatment.

%\textit{Do file: } \texttt{mechanisms.do}
\end{table}

@@

\begin{figure}[H]
     \caption{Experiment description}
    \label{exp_description}
    \begin{center}
    \begin{subfigure}{.50\textwidth}
      \caption{Timeline of the experiment}
        \centering
        \includegraphics[width=\textwidth]{Figuras/timeline_suc_exp_extended.pdf}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
    \caption{Experiment arms}
       \centering
      \includegraphics[width=\textwidth]{Figuras/exp_arms.PNG}
    \end{subfigure}
    \end{center}
         \scriptsize Panel (a) displays in black arrows the period covered by our administrative data: from \hl{XXX} (\hl{XXX} days before the experiment began) to \hl{XXX}. The gray lines arrows correspond to the dates of implementation of the experiment. We finished the experiment earlier in branches 1 and 2 because we realized that we would have enough sample size with the remaining branches and save on operational costs. Panel (b) shows our 5 treatment arms: control (status quo contract), no choice with fee, no choice with promise, choice among both contracts where the commitment contract has a fee, and choice among both contracts where the commitment contract involves only a promise. Within each arm it displays the number of observations. And  for the choice groups it splits the sample size into those who chose the commitment contract (shaded blue triangle) and those that did not (un-shaded triangle).
%\textit{Do file: }  \texttt{timeline\_suc\_exp.do}
\end{figure}

@@

    \begin{subfigure}{0.45\textwidth}
        \caption{Financial cost -- Promise-forcing}
        \centering
        \includegraphics[width=\textwidth]{Figuras/fc_te_pro_3.pdf}
    \end{subfigure}
    \begin{subfigure}{0.45\textwidth}
        \caption{Losing Pawn -- Promise-forcing}
        \centering
        \includegraphics[width=\textwidth]{Figuras/def_te_pro_3.pdf}
    \end{subfigure}
    \end{center}
             \footnotesize \textit{Notes: } 
      \footnotesize{ } \textit{Do file: }  \texttt{fc\_te\_choice\_dec.do, def\_te\_choice\_dec.do, fc\_te.do, def\_te.do }
\end{figure}

@@

\begin{figure}[H]
        \caption{ECDF of Financial Cost}
    \label{ecdf_fc}
    \begin{center}
        \centering
        \includegraphics[width=0.5\textwidth]{Figuras/cdf_fc_pro_2.pdf}
    \end{center}
    \footnotesize \textit{Notes: } 
     \footnotesize{ \textit{Do file: }  \texttt{ecdf\_fc.do}}
\end{figure}

### 2020-05-30 presentations/beamer_pola.tex
update repo structure

\input{preamble_beamer.tex}

@@

\author[]{
Isaac Meza\inst{1}\and Joyce Sadka\inst{2}  \and Enrique Seira\inst{2}}
\institute{ \inst{1} University of California, Berkeley \and
  \inst{2}  Instituto Tecnológico Autónomo de México}

\date{Oct 4, 2019\\\;\;\textsc{Applied Lunch}\\\;\;\textbf{\alert{VERY PRELIMINARY!} }}

@@

\begin{frame}{Motivation}
\begin{itemize}
\vfill \item True story:
    \begin{itemize}
        \item A large Mexican pawn shop (PS) contacted us worried that their default was too high...
        \item ...conjecturing it had to do with contract without frequent payments.
    \end{itemize}
    \pause
\vfill \item Their contract:
    \begin{itemize}
        \item Accepted gold jewelry as collateral, gave you 70\% of the value of gold in cash.
        \item 7\% Monthly interest in outstanding debt.
        \item 3 month term. You could pay principal+interest anytime with no penalty and recover your item. 
        \item Only 33\% paid before day 85th.
        \item if in 90 (+15 days grace) you did not do this, you lose your item.
    \end{itemize}
    \pause
\vfill \item Does forcing them to \textbf{pay monthly} increase payment (decrease default)?
    \begin{itemize}
        \item PS prefers clients recover their item (non-profit mandate).
        \item Higher frequency has costs: $\uparrow$ transaction costs, $\downarrow$  welfare (with neo-classical consumers).
        \item But may help present biased consumers. 
    \end{itemize}
    \vfill
    \pause
    \item \textbf{Is there demand for commitment?}
\end{itemize}
\vfill
\end{frame}

\begin{frame}{Why could frequency decrease default?}
    \begin{itemize}
        \vfill \item Behavioral:
        \begin{itemize}
            \item Commitment device not to spend the money (for present biased consumers).
            \item Creates a habit for saving (or paying) / forces to make a plan (for inattentive consumers).
            \item Sunk cost fallacy (having already paid some).
        \end{itemize}
    \end{itemize}

    \begin{itemize}
        \vfill \item Neoclassical:
        \begin{itemize}    
        \item Increases consumer trust with branch personnel?
        \item Pecuniary/psychological cost from not paying?
        \item Less ``distance to goal''.
         \item Borrows from other lenders to pay $\rightarrow$ more monitoring.
        \end{itemize}
        \vfill
    \item Neoclassical consumers would \textbf{not} tie their hands for free.
        \begin{itemize}
            \item Demand for commitment would signal the presence of present biased consumers.
        \end{itemize}
    \end{itemize}   
    \vfill
\end{frame}

@@

\begin{itemize}
    \item There needs to be a punishment from not paying in order to have commitment.
    \vfill
    \item But punishment if it is too hard, and there are random shocks to income, even present biased consumers will not demand it.
    \vfill 
    \item Two types of punishment in different arms:
    \begin{itemize}
        \item Monetary fee: $4\%$ of missed payment.
        \item Psychological cost of lying: made promise very salient.
    \end{itemize}
\end{itemize}

\begin{frame}{Literature in Microfinance}

        \begin{itemize}
        \vfill \item Field and Pande ``\textit{Repayment Frequency and Default in Microfinance}'' JEEA 2008.
            \begin{itemize}
            \item Randomized 100 MFI groups to weekly vs monthly payment.
            \item Found no effect of frequency (default in about 1\% for all).
            \end{itemize}
            \pause
        \vfill \item Field et al ``Repayment flexibility can reduce financial stress'' PLOS 2012
            \begin{itemize}
            \item Randomized ~100 MFI groups to weekly vs monthly payment.
            \item Found those in weekly payment  50\% more likely to report feeling anxious.
            \end{itemize}
            \pause
        \vfill \item Field, Pande, Papp, Rigol ``Does the classic microfinance model discourage entrepreneurship among the poor?'' AER 2013
        \pause
        \vfill \item Ahsarf et al ``Tying odysseus to the mast'' QJE 2006.
               \begin{itemize}
                \item Offered savings where they could not withdraw for a given time /goal (~700 people, 28\% accepted).
                \item Found present biased women  are more 15pp likely to take up.
                \item 80\% increase in savings 1yr later (ITT).
             \end{itemize}
             \vfill

    \end{itemize}

@@

\begin{figure}[H]
    \label{nochoicehte_takeup_pred}
    \begin{center}
     \begin{subfigure}{0.45\textwidth}
        \caption{In the last six months, did you need money to pay...?}
        \centering
        \includegraphics[width=\textwidth]{Figuras/perc_utilities.pdf}
    \end{subfigure}
     \begin{subfigure}{0.45\textwidth}
        \caption{Cost and Time to arrive to pawn shop}
        \centering
        \includegraphics[width=\textwidth]{Figuras/cost_time_scatter_hist.png}
    \end{subfigure}
    \end{center}
    \end{figure}

@@

\begin{frame}[label = hte]{TE heterogeneity in Recovering}
    \begin{center}
        \includegraphics[width=.80\textwidth]{Figuras/he_dist_des_c_pro_2.pdf}
    \end{center}

@@

\begin{frame}{Some determinants of TE - The nature of heterogeneity}
    \begin{center}
        \includegraphics[width=.9\textwidth]{Figuras/HE/he_int_des_c_pro_2.pdf}
    \end{center}
\end{frame}

@@

@@

\begin{frame}{Those taking up are not those with highest TE}
    \begin{center}
        \includegraphics[width=.70\textwidth]{Figuras/takeup_he_pro_2_fee.pdf}
    \end{center}
\end{frame}

@@

\begin{frame}{Next steps}

\begin{itemize}
    \item Can a simple neoclassical model explain quantitatively the results?
    \vfill
    \item Estimate a model and talk about welfare?
    \vfill
    \item What about naive present-biased consumers? Can we separate them?
    \vfill
    \item Does promise commitment work for some people?
    \vfill
    \item Does frequent payment cause lack of payment of basic services or other loans?
\end{itemize}

@@

@@

\begin{frame}{ML fit}
    \begin{center}
    \scriptsize
    \input{./Tables/oos_pago_frec_vol_fee.tex}
    \end{center}
\end{frame}

@@

where $\psi$ is a known score function and $h$ is an unknown nuisance function also estimated from data.\\

>>>>

\input{presentations/preamble_beamer.tex}

@@

\author[]{
Isaac Meza\inst{1}\and Joyce Sadka\inst{2}  \and Enrique Seira\inst{2}}

\institute{ \inst{1} University of California, Berkeley \and
  \inst{2}  Instituto Tecnológico Autónomo de México}

\date{Oct, 2019\\\;\;\\\;\;}

@@

\begin{frame}{Motivation}
\begin{itemize}
\item How the collaboration started
    \begin{itemize}
        \item We were contacted by the organization about their contract.
        \item Conjectured that frequent payments would lead to higher recovery.

    \end{itemize}
    \pause
\vfill \item The standard contract:
    \begin{itemize}
        \item Accepted gold jewelry as collateral, gave you 70\% of the value of gold in cash.
        \item 7\% Monthly interest in outstanding debt.
        \item \textbf{3 month term}. You could pay principal+interest anytime with no penalty and recover your item. 
        \item if in 90 (+15 days grace) you did not do this, you lose your item.
        \item A bit more than 1/3 lost their pawn.
        \item Only 33\% paid before day 85th in our data/
    \end{itemize}
    \pause
    \end{itemize}

\vfill
\end{frame}

\begin{frame}{2 Questions}
        \begin{enumerate}
 \item Does forcing clients to \textbf{pay monthly} increase payment (decrease default)?
    \begin{itemize}
        \item \textbf{Higher frequency has costs}: $\uparrow$ transaction costs, $\downarrow$  welfare (with neo-classical consumers).
        \item But may help present biased consumers. 
    \end{itemize}

    \vfill
    \pause
    \item \textbf{Is there demand for commitment?}
        \begin{itemize}
            \item Would clients demand it if given a choice between monthly payments vs at anytime in 3 months?
            \item Who demands it?
            \item Neoclassical consumers would \textbf{not} tie their hands for free.
        \end{itemize}
\end{enumerate}
\end{frame}

@@

\begin{itemize}
    \item There needs to be a punishment from not paying in order to have commitment.
    \vfill
    \item But punishment if it is too hard, and there are random shocks to income, even present biased consumers will not demand it.
    \vfill 
    \item Tried with 2 types of punishment:
    \begin{itemize}
        \item Monetary fee: $4\%$ of missed payment.
        \item Psychological cost of lying: made promise very salient.
    \end{itemize}
\end{itemize}
\end{frame}

@@

\begin{figure}[H]
    \label{nochoicehte_takeup_pred}
    \begin{center}
     \begin{subfigure}{0.40\textwidth}
        \caption{In the last six months, did you need money to pay...?}
        \centering
        \includegraphics[width=\textwidth]{Figuras/perc_utilities.pdf}
    \end{subfigure}
     \begin{subfigure}{0.5\textwidth}
        \caption{Cost and Time to arrive to pawn shop}
        \centering
        \includegraphics[width=\textwidth]{Figuras/cost_time_scatter_hist.png}
    \end{subfigure}
    \end{center}
    \end{figure}

@@

\begin{frame}[label = hte]{TE heterogeneity in Recovering}
    \begin{center}
        \includegraphics[width=.80\textwidth]{Figuras/he_dist_des_c.pdf}
    \end{center}

@@

\begin{frame}{A proxy for financial cost to client}

Financial cost is defined as: sum of payments + cost of piece if not recovered

    \begin{center}
        \includegraphics[width=.7\textwidth]{Figuras/hist_fc.pdf}
    \end{center}
\end{frame}

\begin{frame}{Effect of Frequency on Financial Cost}

\begin{itemize}
    \item Financial cost decreases by 350 pesos, from a mean of 3,200.
\end{itemize}
\vspace{.2in}
\vfill
    \begin{center}
        \includegraphics[width=.9\textwidth]{Figuras/FC_table.png}
    \end{center}
\end{frame}

@@

\begin{frame}{Fit: Machine learning models}
\begin{itemize}
    \item 87\% accuracy: we can predict take up well.
    \vfill
\end{itemize}
    \begin{center}
    \input{./Tables/oos_pago_frec_vol_fee.tex}
    \end{center}
\end{frame}

@@

\begin{frame}{``Benefit'' positively correlated with predicted take-up}

\begin{itemize}
    \item Estimate take-up \alert{in Choice fee group} as a function of X's: $f(\beta,X)$.
    \item Predict take up using $f(\beta,X)$ in \alert{No-choice group}.
    \item relate predicted take up to effect they really had in \alert{No-choice group}.
\end{itemize}

\begin{figure}[H]
    \begin{center}
     \begin{subfigure}{0.45\textwidth}
        \caption{Recovery vs predicted take-up}
        \centering
        \includegraphics[width=\textwidth]{Figuras/takeup_he_pro_2_pago_frec_vol_fee_des_c.pdf}
    \end{subfigure}
     \begin{subfigure}{0.45\textwidth}
        \caption{FC vs predicted take-up}
        \centering
        \includegraphics[width=\textwidth]{Figuras/takeup_he_pro_2_pago_frec_vol_fee_fc_survey.pdf}
    \end{subfigure}
    \end{center}
    \end{figure}

@@

\begin{frame}{Potential projects and remaining questions}

\begin{enumerate}

    \item Current project:
    \begin{enumerate}
        \item Get historical data to see results by different types of people.
        \item Replicate experiment in 2 or 3 branches, with improved design.
        \begin{itemize}
            \item Ask preference to all, then randomize.
            \item Better satisfaction survey.
            \item Measure preferences better at baseline, and expectations.
            \item Cross check with bureau and savings.
            \item Measure mistakes in take up
        \end{itemize}
    \end{enumerate}

    \vfill
    \item Impact of different collateral requirements:
    \begin{itemize}
        \item Pilot 65\% and 75\% collateral and measure take up, recovery and profits
    \end{itemize}

    \vfill
    \item Reporting to Credit Bureau (pilot):
    \begin{itemize}
        \item As a way to impose discipline (ex-ante vs ex-post)
        \item It can be voluntary
    \end{itemize}

    \vfill
    \item Optimal auctions to sell the goods in a webpage.

    \vfill
    \item Tracking by cellphone to screen for bank.

\end{enumerate}
\end{frame}

@@

\begin{frame}{Some determinants of TE - The nature of heterogeneity}
    \begin{center}
        \includegraphics[width=.9\textwidth]{Figuras/he_int_vertical_des_c_pro_2.pdf}
    \end{center}
\end{frame}

@@

@@

where $\psi$ is a known score function and $h$ is an unkown nuisance function also estimated from data.\\
